---
title: "Danish Kings"
author: "Anonymous"
date: "2026-12-03"
output: html_document
---

#What function did you use to load your dataset in R as a tibble? What do you see when you run head() on your ‘kings’ digital object? 
```{r}
library(tidyverse)

monarchs <- read_csv2("data/Danish Monarchs.csv")
```
```{r}
glimpse(monarchs)
```


```{r}
head(monarchs)
```

#What is the average age of reign among your monarchs and how many rulers have exceeded it? 
```{r}
monarchs <- monarchs %>%
  mutate(reign_length = Regeringsslut - Regeringsstart)

average_reign <- mean(monarchs$reign_length, na.rm = TRUE)

above_average <- monarchs %>%
  filter(reign_length > average_reign)


cat("Average reign length:", average_reign, "years\n")
cat("Number of rulers above average:", nrow(above_average))
```


```{r}
longest_monarchs <- monarchs %>%
  mutate(
    reign_length = Regeringsslut - Regeringsstart,
    reign_days = reign_length * 365
  ) %>%
  arrange(desc(reign_length)) %>%
  slice(1:3)

longest_monarchs %>%
  select(Navn, reign_length, reign_days)
```


#How many days did the three longest-ruling monarchs rule? 

```{r}
longest_monarchs <- monarchs %>%
  mutate(
    reign_length = Regeringsslut - Regeringsstart
  ) %>%
  arrange(desc(reign_length)) %>%
  slice(1:3)

longest_monarchs %>%
  select(Navn, reign_length)
```

#Plot the duration of reign in time and briefly answer “What is the long-term trend in the duration of reign among monarchs?” 

```{r}
monarchs_plot <- monarchs %>%
  mutate(
    reign_length = Regeringsslut - Regeringsstart,
    mid_year = (Regeringsstart + Regeringsslut) / 2
  )

ggplot(monarchs_plot, aes(x = mid_year, y = reign_length)) +
  geom_point(color = "blue") +
  geom_smooth() +
  theme_minimal() +
  labs(
    title = "Reign Length of Danish Monarchs Over Time",
    x = "Year",
    y = "Reign Length (Years)"
  )
```










