---
title: "kvinder_1860_analysis"
author: "anonymous"
date: "2026-05-07"
output: html_document
---
The data is processed with tidyverse and ggplot

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

data <- read.csv(
  "data/census-1860-original-csv-(3).csv",
  fileEncoding = "UTF-8"
)
```

# Civilstand of the women

I am using a ggplot to create a graph that illustrates women's marital status.

```{r}
ggplot(data, aes(x = civilstand)) +
  geom_bar(fill = "red") +
  labs(
    title = "Kvinders civilstand i Aarhus 1860",
    x = "Civilstand",
    y = "Antal kvinder"
  )
```


# Married vs unmarried women

```{r}
civil_subset <- data %>%
  filter(civilstand %in% c("gift", "ugift"))

ggplot(civil_subset, aes(x = civilstand)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Gift vs ugift kvinder i Aarhus 1860",
    x = "Civilstand",
    y = "Antal kvinder"
  )
```


# Top 10 most frequent occupations for woman 
I am making a table of the most frequent occupations for women in Aarhus in the 1860's as a preparation for a ggplot. 

```{r}
top10 <- data %>%
  count(stilling, sort = TRUE) %>%
  slice(1:10)

top10
```


# Top 10 most frequent occupations for woman

Here is the ggplot that illustrates the frequency itself

```{r}
ggplot(top10, aes(x = reorder(stilling, n), y = n)) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(
    title = "Top 10 kvindelige stillinger i Aarhus 1860",
    x = "Stilling",
    y = "Antal kvinder"
  )
```



# Occupations among married and unmarried women

```{r}
top10_stillinger <- data %>%
  count(stilling, sort = TRUE) %>%
  slice(1:10)

civil_stilling <- data %>%
  filter(
    civilstand %in% c("gift", "ugift"),
    stilling %in% top10_stillinger$stilling
  ) %>%
  count(civilstand, stilling)

ggplot(civil_stilling,
       aes(x = reorder(stilling, n), y = n, fill = civilstand)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Top 10 stillinger blandt gifte og ugifte kvinder",
    x = "Stilling",
    y = "Antal kvinder",
    fill = "Civilstand"
  )
```

